### 2017-11-23 supplementary_figures_nar.tex
Edited legends of supplementary figures.

\PassOptionsToPackage{utf8}{inputenc}
\documentclass[12pt]{report}

\usepackage[margin=2cm]{geometry}

\usepackage{graphicx}

\renewcommand{\thefigure}{S\arabic{figure}}

\begin{document}

\begin{figure}[b]
	\centerline{\includegraphics[width=\textwidth]{nar_figures/supp_figure_1.eps}}
  \caption{ Pearson correlation between contact profiles and independent
copy number estimation from the COSMIC data base (Forbes \textit{et al}.,
2010). A. T47D breast cancer cell line (6 samples). B. K562 leukemia cell
line (8 samples). Each dot represents a sample normalized by a different
tool. The data normalized by \textit{OneD} correlates better with the copy
number than the unnormalized data.}
\end{figure}

\begin{figure}
	\centerline{\includegraphics[width=\textwidth]{nar_figures/supp_figure_2.eps}}
  \caption{ Distribution of the pair-wise comparisons between the samples
included in the T47D set. X axis: similarity metric. Y axis: correction
method. Each panel groups pairs of samples with the corresponding
characteristics (in terms of cell type and protocol). A. Pearson
correlation between observed over expected counts. B. Spearman correlation
between observed counts. C. Stratum-adjusted correlation coefficient (SCC)
between observed counts. D. Reproducibility score of observed counts.}
\end{figure}

\begin{figure}
	\centerline{\includegraphics[width=\textwidth]{nar_figures/supp_figure_3.eps}}
  \caption{Distribution of the pair-wise comparisons between the
samples included in the K562 set. X axis: similarity metric. Y axis:
correction method. Each panel groups pairs of samples with the
corresponding characteristics (in terms of cell type and protocol). A.
Pearson correlation between observed over expected counts. B. Spearman
correlation between observed counts. C. Stratum-adjusted correlation
coefficient (SCC) between observed counts. D. Reproducibility score of
observed counts.} \end{figure}

\begin{figure}
	\centerline{\includegraphics[width=\textwidth]{nar_figures/supp_figure_4.eps}}
    \caption{Removing biases from Hi-C on aberrant karyotypes (Spearman
correlation). A and B. Average changes compared to the raw data on the
T47D and K562 data sets. The bars represent 95\% confidence intervals
centered on the mean difference of the Spearman correlation between a
given correction method and the raw data. C and D. ROC curves on the T47D
and K562 sets. The areas under the curve are indicated in the bottom right
corner. The color code is the same as in panels A and B.}
\end{figure}

\begin{figure}
	\centerline{\includegraphics[width=\textwidth]{nar_figures/supp_figure_5.eps}}
  \caption{Removing biases from Hi-C on aberrant karyotypes (Pearson
correlation). A and B. Average changes compared to the raw data on the
T47D and K562 data sets. The bars represent 95\% confidence intervals
centered on the mean difference of the Pearson correlation of the observed
over expected counts between a given correction method and the raw data. C
and D. ROC curves on the T47D and K562 sets. The areas under the curve are
indicated in the bottom right corner. The color code is the same as in
panels A and B.}
\end{figure}

\begin{figure}
	\centerline{\includegraphics[width=\textwidth]{nar_figures/supp_figure_6.eps}}
  \caption{Removing biases from Hi-C on aberrant karyotypes (SCC). A and
B. Average changes compared to the raw data on the T47D and K562 data
sets. The bars represent 95\% confidence intervals centered on the mean
difference of the stratum-adjusted correlation coefficient (SCC) between a
given correction method and the raw data. C and D. ROC curves on the T47D
and K562 sets. The areas under the curve are indicated in the bottom right
corner. The color code is the same as in panels A and B.}
\end{figure}

\begin{figure}
	\centerline{\includegraphics[width=\textwidth]{nar_figures/supp_figure_7.eps}}
  \caption{Distribution of the pair-wise comparisons between the samples
included in the mm10 data set. X axis: similarity metric. Y axis:
correction method. Each panel groups pairs of samples with the
corresponding characteristics (in terms of cell type and protocol). A.
Pearson correlation between observed over expected counts. B. Spearman
correlation between observed counts. C. Stratum-adjusted correlation
coefficient (SCC) between observed counts. D. Reproducibility score of
observed counts.}
\end{figure}

\begin{figure}
	\centerline{\includegraphics[width=\textwidth]{nar_figures/supp_figure_8.eps}}
  \caption{Removing biases from Hi-C on normal karyotypes (Spearman
correlation). A. Average changes compared to the raw data on the mm10 data
set. The bars represent 95\% confidence intervals centered on the mean
difference of the Spearman correlation between a given correction method
and the raw data. B. ROC curves on the mm10 data set. The areas under the
curve are indicated in the bottom right corner. The color code is the same
as in panel A.}
\end{figure}

\begin{figure}
	\centerline{\includegraphics[width=\textwidth]{nar_figures/supp_figure_9.eps}}
  \caption{Removing biases from Hi-C on normal karyotypes (Pearson
correlation). A. Average changes compared to the raw data on the mm10 data
set. The bars represent 95\% confidence intervals centered on the mean
difference of the Pearson correlation of the observed over expected counts
between a given correction method and the raw data. B. ROC curves on the
mm10 data set. The areas under the curve are indicated in the bottom right
corner. The color code is the same as in panel A.}
\end{figure}

\begin{figure}
	\centerline{\includegraphics[width=\textwidth]{nar_figures/supp_figure_10.eps}}
  \caption{Removing biases from Hi-C on normal karyotypes (SCC). A.
Average changes compared to the raw data on the mm10 data set. The bars
represent 95\% confidence intervals centered on the mean difference of the
stratum-adjusted correlation coefficient (SCC) between a given correction
method and the raw data. B. ROC curves on the mm10 data set. The areas
under the curve are indicated in the bottom right corner. The color code
is the same as in panel A.}
\end{figure}

\begin{figure}
	\centerline{\includegraphics[width=\textwidth]{nar_figures/supp_figure_11.eps}}
    \caption{
    Removing biases from Hi-C on GM12878 samples. X axis: reproducibility
    score. Y axis: correction method. The bars represent 95\% confidence intervals centered on the mean of the
    reproducibility score of each correction method. The vertical dark purple
    line marks the reproducibility score for the uncorrected data (raw). Upper panel: pairs of
    samples processed using different protocols. Lower panel: pairs of samples
    processed using the same protocol.}
\end{figure}

\begin{figure}
	\centerline{\includegraphics[width=\textwidth]{nar_figures/supp_figure_12.eps}}
    \caption{
    Distribution of the simulation results. X axis: Difference in
    reproducibility score between pairs of samples of the same cell type and
    pairs of samples of different cell type. Y axis: correction method}
\end{figure}

\begin{figure}
	\centerline{\includegraphics[width=\textwidth]{nar_figures/supp_figure_13.eps}}
    \caption{Correlation between subsamples at different resolutions and
    depths. X axis: sequencing depth. Y axis: Pearson correlation between the
    correction vectors of two different subsamples. Each color represents a
    correction method and each panel a bining resolution. Note the Y axis limits
    are specific to each panel.}
\end{figure}

\end{document}